---
title: "Northern Latitudes - Exploratory Data Analysis"
subtitle: "Project STRETCH"
author: "Dana K Briscoe"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: none
    df_print: paged
    highlight: tango
    number_sections: no
    theme: flatly
    toc: yes
    toc_float: true
    toc_collapsed: true
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{=html}
<style>
.main-container {
    max-width: 1400px !important;
    margin-left: 0;
    margin-right: 5;
}

</style>

<style type="text/css">
  body{
  font-size: 12.5pt;
}
</style>
```
<!-- Load rmd files -->

```{r 01, child=c('01_northern_lats_EDA_maps.Rmd')}

```

```{r 02, child=c('02_northern_lats_EDA_cc_plots.Rmd')}

```

```{r 03, child=c('03_northern_lats_EDA_cc_env_figs.Rmd')}

```

```{r 04, child=c('04_northern_lats_EDA_tzcf_figs.Rmd')}

```

```{r 05, child=c('05_northern_lats_EDA_gam.Rmd')}

```

<br>

## 1 \| Env-Cohort 1

------------------------------------------------------------------------

<span style="font-size:14pt">

### Fig 1. Turtles & SSTa {.tabset .tabset-pills}

Figure 1 shows the average monthly SST anomalies (SSTa) from July - Aug, 2023. Turtles deployed in July 2023 (n=25) are referred to as 'Cohort 1'. The tracks from 11 July - 20 Sept 2023 are shown in colors corresponding to the Wildlife Computers [map](https://my.wildlifecomputers.com/data/map/?id=646c14fae9b3514a9430a544). <br>

The 1.5°C contour is shown in black. The Thermal Corridor study area is outlined in gray. SSTa source: [NOAA Coral Reef Watch, Monthly 5km](https://coralreefwatch.noaa.gov/product/5km/index_5km_sst.php).

The tabs can be clicked to change the spatial extent. </span>

#### Full Extent

```{r fig-1-ssta-cohort-1-map-full,out.height="110%"}
p_ssta_mean_tcms_w_contour

```

#### Zoomed

```{r fig-1-ssta-cohort-1-map-zoom, out.height="90%"}
p_ssta_mean_tcms_w_contour_zoom
```

### Fig 2. Turtles & Chla {.tabset .tabset-pills}

Figure 2 shows the average monthly chlorophyll-a (Chla) from July - Aug, 2023 and cohort 1 turtle tracks (n=25, July-Sept 2023). <br>

The Thermal Corridor study area is outlined in gray. Chla source: [NOAA S-NPP VIIRS, Monthly 4km](https://coastwatch.pfeg.noaa.gov/erddap/info/nesdisVHNSQchlaMonthly/index.html).

The tabs can be clicked to change the spatial extent.

#### Full Extent

```{r fig-1-chl-cohort-1-map-full, out.height="120%"}
p_chla_mean_tcms
```

#### Zoomed

```{r fig-1-chl-cohort-1-map-zoom, out.height="90%"}
p_chla_mean_tcms_zoom
```

<br>

## 2 \| Max Lats-Cohort 1

------------------------------------------------------------------------

### Table 1. Max Lats by Turtle ID

```{r table-1-max-lats-cohort-1, echo=FALSE}
daily_avg_data_cohort_1_table %>%
   # relocate(c(`Max Lat`, Date), .before = ID) %>%
kbl(caption = " Northernmost latitude reached by each individual turtle, ranked in descending order.") %>%
    kable_paper("hover", full_width = T) %>%
    kable_styling(bootstrap_options = c("condensed", "responsive")) %>%
    kable_styling(fixed_thead = T) %>%
    kable_minimal() %>%
scroll_box(width = "100%", height = "300px")
```

<br>

### Fig 3. Cohort 1 Max Lats by Date

Figure 3 shows the northern progression of each turtle from Cohort 1, since their release date (11 July 2023).

Use the mouse to show tag information for each line.

```{r fig-daily-avg-cohort-1-max-lats-date, out.height="80%", out.width="80%"}
gg_daily_avg_cohort1_max_lats
```

<br> <br>

### Fig 4. Historic vs Cohort Max Lats

Figure 4 compares the average latitudes experienced by 2023's Cohort 1 turtles (green line) and turtles from the historic data set ('Historic' 1997-2013, purple line). Latitudes were spatially averaged eastern North Pacific region, between 160°W to 130°W longitude. The gray bands represent ± 2 standard deviations.

Use the mouse to show tag information for each line.

```{r fig-gg-avg-cc-lats-by-week, out.height="80%", out.width="90%"}
gg_p_avg_turtle_lats_by_week
```

<br>

## 3 \| TZCF Locations

------------------------------------------------------------------------

### Fig 5. Seasonal TZCF

Figure 5 shows the seasonal position of the 18°C SST isotherm between 160°W and 130°W longitude. This isotherm is a common proxy for the Transition Zone Chlorophyll Front.

For each year between 1997-2023 (which encompasses both the historic tags and current, 'Cohort 1'), the annual isotherm trend is plotted as a gray line. The years 2015, 2019, 2020 (light reds) and 2023 (dark red) are highlighted as they were anomalously warm years.

```{r fig-gg-tzcf-all-yrs, out.height="80%", out.width="95%"}
gg_tzcf_all_yrs

```

<br>

### Fig 6. TZCF & Turtles

The average latitudinal movements of loggerheads from the Historic data set (purple line, 1997-2013, n=69) and Cohort 1 (green line, 2023, n=25) between 160°W and 130°W are shown in relation to the 18°C SST isotherm. The all-time monthly average of isotherm position is shown in gray (1997-2022); the 2023 monthly position is shown in red.

```{r fig-gg_tzcf_hist_cohort, out.height="80%", out.width="95%"}
gg_tzcf_hist_cohort

```

## 4 \| Historic Summary Stats

------------------------------------------------------------------------

### Table 2. n Tags 160°W-130°W

Between 1997-2013, there were `r length(unique(hist_df_subset$id))` turtles that transmitted between 160W and 130W. Of these, `r length(unique(hist_df_subset_septs$id))` transmitted during the month of September.

<br>

The table below shows the number of turtles that transmitted data (160°W-130°W), across individual months and years (1997-2013). Note that a few individuals transmitted for multiple years, which is reflected in the monthly column totals.

<!-- ### Table 2. n Tags per Year, Month -->

```{r hist-freq-tags-per-year-month}
ntags_hist_tags_160W_130W_per_year_month 

```

<!-- ### Fig 7. Historic Tags in Sept -->

<!-- Of the 69 tags (160W-130W), this figure shows the length of each track over time, and which tags transmitted during the month of September (n=31 out of 69).  -->

<!-- September transmissions for each individual track are shown in red. -->

<!-- ```{r fig-gg-graph-hist-id-by-yrmon, out.height="150%", out.width="100%"} -->

<!-- gg_graph_hist_id_by_yrmon -->

<!-- ``` -->

<br>

### Fig 7. Sept Lats: Mean vs Median

Summary statistics for the historic data set show there is a negative skew to the distribution of September latitudes. <br> **The mean of September latitudes: `r hist_all_stats$value[1]`** <br> **The median of September latitudes: `r hist_all_stats$value[2]`**

```{r}
p_hist_df_avg_vs_med
```

### 

## 5 \| GAM Results

------------------------------------------------------------------------

The following GAM model explores evidence that turtle latitude has increased over time.

For the Historic & Cohort 1 data sets, the median latitude and longitudes were calculated for each year that data was available.

The table below shows the input data:

<br>

### Table 3. Median September Turtle Latitude & Longitude By Year

Table shows the median September latitude and longitudes for turtles (160°W-130°W).

```{r}
max_lat_yr_w2023 %>%
    mutate(med_lat = round(med_lat,2),
            med_lon = round(med_lon,2)) %>%
    dplyr::select(c(year, med_lat, med_lon)) %>%
    rename("Year"="year", "Median \nLatitude (°N)"="med_lat", "Median \nLongitude (°W)" = "med_lon")
    
```

<br>

### Model Outputs

The following models were run (with and without the year 2023):

<br>

**Median September Latitude** \~ **Median September Longitude** + **Year**

<br>

### Fig 8. Model Summary (with 2023)

```{r summary-gam_model_w2023, out.height="40%", out.width="95%"}
summary(gam_model_w2023)

draw(gam_model_w2023, residuals = TRUE)

```

------------------------------------------------------------------------

### Fig. 9 Model Summary (without 2023)

```{r summary-gam_model_wo-2023, out.height="40%", out.width="95%"}
summary(gam_model_wo_2023)

draw(gam_model_wo_2023, residuals = TRUE)
```
